Likelihood-Based Inference for Generalized Linear Mixed Models: Inference with Package \pkg{glmm}

### 2017-02-14 JSS/article.tex
started to edit section on formatting data

\subsubsection{The salamander data}

@@

The following vectors can be used to fit a generalized linear mixed model using the \texttt{glmm} package. These vectors can be contained in a data frame, but they do not need to be.
\begin{enumerate}
\item A response vector. If your response is Poisson, then the entries in the response vector must be natural numbers. If your response is Bernoulli, then the entries in the response vector must be $0$ and $1$. (For this version of \texttt{glmm}, these are the only two response types possible. If you need to fit a model with a different response, contact me.) 
\item At least one vector that will be used for defining the random effects' design matrix. For this version of \texttt{glmm}, the vector(s) should be class \texttt{factor}.
\item Vector(s) that will be used for defining the fixed effects' design matrix. The vector(s) can be of class \texttt{factor} or \texttt{numeric}. 
\end{enumerate}
The first two types of vectors described in the list are required. The last type is optional. That is, the minimum requirement to fit a \texttt{glmm} model is the response vector and one vector for defining the random effects' design matrix. \\

We use the \texttt{salamander} dataset as an example in this vignette. For your convenience, it is already included in the \texttt{glmm} package. The data arose from an experiment conducted at the University of Chicago in 1986 and were first presented by \citet[section 14.5]{mcc:nelder:1989}. Scientists paired female and male salamanders of two types (Rough Butt and White Side) and collected data on whether or not they mated.  \\

@@

The first R  command shown below gives us access to the \texttt{glmm} package and and all of its commands. The second line of code gives us access to the \texttt{salamander} data frame.  The next three commands help us begin to understand the data. We have four variables: \texttt{Mate}, \texttt{Cross}, \texttt{Female}, and \texttt{Male}. The summary shows us \texttt{Mate} is numeric, \texttt{Cross} is a factor with four levels, \texttt{Female} is a factor, and \texttt{Male} is a factor.

>>>>

\subsubsection{The salamander data}\label{sec:salex}

@@

The following vectors can be used to fit a generalized linear mixed model using the \texttt{glmm} package. These vectors can be contained in a data frame, but they do not need to be.
\begin{enumerate}
\item The response vector(s). If your response is Poisson or Bernoulli, then you should have a single response vector. If your response is Poisson, then the entries in the response vector must be natural numbers. If your response is Bernoulli, then the entries in the response vector must be $0$ and $1$. If your response is binomial, then you should have two vectors connected with \texttt{cbind}: the first vector should be the number of successes and the second should be the number of failures.
\item At least one vector that will be used for defining the random effects' design matrix. For this version of \texttt{glmm}, the vector(s) should be class \texttt{factor}.
\item Vector(s) that will be used for defining the fixed effects' design matrix. The vector(s) can be of class \texttt{factor} or \texttt{numeric}. 
\end{enumerate}
The first two types of vectors described in the list are required. The last type is optional. That is, the minimum requirement to fit a \texttt{glmm} model is the response vector and one vector for defining the random effects' design matrix. \\

\subsection{Salamander data analysis}

Recall the salamander experiment described in Section \ref{sec:examples}. For your convenience, the \texttt{salamander} dataset is  included in the \texttt{glmm} package. The data arose from an experiment conducted at the University of Chicago in 1986 and were first presented by \citet[section 14.5]{mcc:nelder:1989}.  \\

@@

The first \proglang{R}  command shown below gives us access to the \texttt{glmm} package and and all of its commands. The second line of code gives us access to the \texttt{salamander} data frame.  The next three commands help us begin to understand the data. We have four variables: \texttt{Mate}, \texttt{Cross}, \texttt{Female}, and \texttt{Male}. The summary shows us \texttt{Mate} is numeric, \texttt{Cross} is a factor with four levels, \texttt{Female} is a factor, and \texttt{Male} is a factor.
